---
title: "Final Project"
output: html_document
date: "2024-01-17"
---

```{r}
library(mgcv)        # GAMs
library(rsample)     # data splitting 
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)       # bagging
```


```{r}
df <- read.csv("phones_data.csv", header=T)

summary(df)
```

```{r}
head(df)
```


```{r warning=FALSE}
# Categorical
df$brand_name <- factor(df$brand_name)
df$os <- factor(df$os)
# To numeric
df$popularity <- as.numeric(df$popularity)

# To euro
df$best_price <- df$best_price*0.024
df$lowest_price <- df$lowest_price*0.024
df$highest_price <- df$highest_price*0.024
```

```{r warning=FALSE}
summary(df)
```

We can notice that there is an outlier in the sellers amount. 

```{r fig.height=20, fig.width=20}
pairs(df[c(-1, -2, -12)], pch=19, lower.panel=panel.smooth)
```

Identify the outlier, here we can see that the row has odd values, for example the best price was 28€ and the lowest price was higher than the highest price. Therefore the screen size is 75 inches.

```{r}
df[which.max(df$screen_size), ]
```

We decided to remove the outlier.

```{r}
df <- df[-which.max(df$screen_size), ]
```

We can notice that now that the outlier is removed the plots are more readable.

```{r fig.height=20, fig.width=20}
pairs(df[c(-1, -2, -12)], pch=19, lower.panel=panel.smooth)
```

```{r}
# Lowest and highest price are missing together

# or drop them
# or impute by mean
# or impute by median
# or impute by regression

df[which(is.na(df$lowest_price)), ]
```

# drop them

```{r}
drop_df <- df[-which(is.na(df$lowest_price)), ]


df_split_drop <- initial_split(drop_df, prop = .7)
df_train_drop <- training(df_split_drop)
df_test_drop  <- testing(df_split_drop)

drop_fit <- lm(best_price ~ popularity + screen_size + battery_size + memory_size + lowest_price, data=df_train_drop)
summary(drop_fit)
```

# impute by mean

```{r}
mean_df <- df
mean_df[which(is.na(df$lowest_price)), 6] <- mean(df[-which(is.na(df$lowest_price)), 6])


df_split_mean <- initial_split(mean_df, prop = .7)
df_train_mean <- training(df_split_mean)
df_test_mean  <- testing(df_split_mean)

mean_fit <- lm(best_price ~ popularity + screen_size + battery_size + memory_size + lowest_price, data=df_train_mean)
summary(mean_fit)
```

# impute by median

```{r}
median_df <- df
median_df[which(is.na(df$lowest_price)), 6] <- median(df[-which(is.na(df$lowest_price)), 6])


df_split_median <- initial_split(median_df, prop = .7)
df_train_median <- training(df_split_median)
df_test_median <- testing(df_split_median)

median_fit <- lm(best_price ~ popularity + screen_size + battery_size + memory_size + lowest_price, data=df_split_median)
summary(median_fit)
```

# impute by regression

```{r}
test <- df[which(is.na(df$lowest_price)), ]
train <- df[-which(is.na(df$lowest_price)), ]

# Impute missing values
fit <- lm(lowest_price ~ popularity + screen_size + battery_size + memory_size, data=train)
imputed <- predict(fit, test)
summary(fit)
```

```{r}
summary(df$lowest_price)
```

```{r}
regression_df <- df
regression_df$lowest_price[which(is.na(df$lowest_price))] <- imputed
summary(regression_df$lowest_price)
```
```{r}

df_split_regression <- initial_split(regression_df, prop = .7)
df_train_regression <- training(df_split_regression)
df_test_regression <- testing(df_split_regression)

imputed_fit <- lm(best_price ~ popularity + screen_size + battery_size + memory_size + lowest_price, data=df_train_regression)
summary(imputed_fit)
```

# Comparison

```{r}
AIC(drop_fit, mean_fit, median_fit, imputed_fit)
```

```{r}
R2 <- c(summary(drop_fit)$adj.r.squared, summary(mean_fit)$adj.r.squared, summary(median_fit)$adj.r.squared, summary(imputed_fit)$adj.r.squared)
names(R2) <- c("drop", "mean", "median", "imputed")

as.data.frame(R2)
```

```{r}
RMSE_fun <- function(model, test) {
  idxs.na <- apply(is.na(test), 1, any)
  names(idxs.na) <- NULL
  test.drop <- test[!idxs.na, ]
  pred <- predict(model, test.drop)
  sqrt(mean((test$best_price - pred)^2))
}
```


```{r}

RMSE <- c(RMSE_fun(drop_fit, df_test_drop), RMSE_fun(mean_fit, df_test_mean), RMSE_fun(median_fit, df_test_median), RMSE_fun(imputed_fit, df_test_regression))
names(RMSE) <- c("drop", "mean", "median", "imputed")

as.data.frame(RMSE)
```



```{r}
df_split <- initial_split(drop_df, prop = .7)
df_train <- training(df_split)
df_test  <- testing(df_split)

clf <- gam(best_price ~ popularity + s(screen_size) + s(battery_size) + s(memory_size) + s(lowest_price) + os, data=drop_df)
summary(clf)
```

```{r}
set.seed(123)
df_split <- initial_split(drop_df, prop = .7)
df_train <- training(df_split)
df_test  <- testing(df_split)
```

```{r}
tree <- rpart(best_price ~ popularity + screen_size + battery_size + memory_size + lowest_price + os, data=df_train)
rpart.plot(tree)
```

```{r}
tree_pred <- predict(tree, df_test)
sqrt(sum((df_test$best_price - tree_pred)^2))
```




```{r}
for (i in 6:ncol(df)-1) {
  hist(df[, i], main=names(df)[i], xlab=names(df)[i])
}
```

```{r}
for (i in 6:ncol(df)-1) {
  hist(log(df[, i]), main=names(df)[i], xlab=names(df)[i])
}
```
