---
title: "Final_Project"
author: "Nicola Cortinovis, Roberta Lamberti, Marta Lucas, Stefano Tumino"
date: "`r Sys.Date()`"
output: html_document
---
# Initial setup
```{r}
#install.packages("mgcv")
#install.packages("dplyr")
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("ipred")
#install.packages("caret")
#install.packages("stringr")
#install.packages("ggcorrplot")
```

## Import libraries
```{r}
library(stringr)
library(mgcv)        # GAMs
library(rsample)     # data splitting
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)
library(ggcorrplot)
```

## Import dataset
```{r}
df <- read.csv("phones_data.csv", header=T)

summary(df)
```

## Variable cleaning

### Remove useless variables
TODO: elenco variabili categoriali e spiegazioni sul dataset
```{r}
# rimozione colonna degli indici
df$X <- NULL
```

### Factorize the categorical variables
```{r}
df$brand_name <- factor(df$brand_name)
df$os <- factor(df$os)
```

 
Let's divide the popularity into levels,assigning three levels associated with a numeric score.
- From 0 to 400 -> low
- From 401 to 800 -> medium
- From 801 to 1224 -> high
```{r}
df$popularity <- as.numeric(df$popularity)

tag <- c(0,400, 800, Inf)

df$popularity_levels <- cut(df$popularity, breaks = tag,
labels=c("low", "medium", "high"), include.lowest=TRUE)
```

### Convert the prices to euros
```{r}

df$best_price <- df$best_price*0.024
df$lowest_price <- df$lowest_price*0.024
df$highest_price <- df$highest_price*0.024
```

### Split the timestamps
```{r}
df$month <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][1]}))
df$year <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][2]}))-2000
```

# Dataset cleaning 
- sistema operativo mancante modelli molto vecchi di telefoni, non sono smartphone 
- memory size mancante per gli stessi 

## Remove duplicated rows
```{r}

idxs <- which(duplicated(df[,c(-2, -4)]))

succ <- idxs[-length(idxs)] - idxs[-1] == -1
succ <- c(succ, FALSE)

i = 1
while (i <= length(idxs)){
  start = idxs[i]-1
  sum <- c(df$popularity[idxs[i]])
  while (succ[i] == TRUE){
    i = i + 1
    sum <- c(sum, df$popularity[idxs[i]])
  }
  df$popularity[start] <- mean(c(sum, df$popularity[start]))
  i = i + 1
}

df <- df[-idxs, ]

```

## Manage missing values
Since there are only 10 missing values for battery size and 2 NAs for screen size, we can drop those observations.

We notice that most of the missing values for memory size are relative to phones that are not smart-phones, for this reason we decided to substitute them with the median of the values for memory size of non-smartphones.TODO: sistemare
For the highest/lowest price we'll just substitute the mean of those columns.

### Drop Battery and Screen size missing values rows
```{r}
df <- df[- which(is.na(df$battery_size)),]
df <- df[- which(is.na(df$screen_size)),]
```

### Impute memory size missing values

We notice that most of the missing values are not smartphones... (os) TODO: spiegare quello che abbiamo fatto per la media
```{r}
tmp <- df[which(df$os == ""),]$memory_size
df$memory_size[which(is.na(df$memory_size))] <- median(tmp[-which(is.na(tmp))])
```


### substitute the highest/lowest price NA with the median of the column
```{r}
tmp <- which(is.na(df$lowest_price))
df$lowest_price[tmp] <- median(df$lowest_price[-tmp])
```

```{r}
tmp <- which(is.na(df$highest_price))
df$highest_price[tmp] <- median(df$highest_price[-tmp])

```

### Remove useless variables
```{r}
# remove useless variables
#df$model_name <- NULL
#df$release_date <- NULL
#df$popularity <- NULL
```

# Data Analysis
## Correlation matrix
```{r}
corr <- cor(df[, c(-1,-2, -3, -12, -13, -14, -15)], use="complete.obs" )

ggcorrplot(corr, hc.order = TRUE, lab = TRUE, colors=c("yellow", "white", "purple"))
```
TODO: explain what columns need to be removed (highest/lowest price)
## Single plots and Outlier analysis
### Numerical
TODO: -boxplots/other ideas
```{r}
cols <- 4:11

for (i in cols) {
  layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
  par(mar=c(0,3.1,1.1,2.1))
  boxplot(df[, i], horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main=colnames(df)[i])
  par(mar=c(4,3.1,1.1,2.1))
  hist(df[, i],border = F, main = "", col="purple")
}
```
### Categorical
```{r}
for (i in c(1, 3, 13)) {
  barplot(sort(table(df[, i])), main=colnames(df)[i], col="purple")
}

for (i in c(1, 3, 13)) {
  boxplot(df$best_price ~ df[, i],  xaxt = "n", col= "yellow", frame= F, main=colnames(df)[i])
}
```

### Pairs plot
```{r fig.height=7.5, fig.width=7.5}
pairs(df[cols], pch=19, lower.panel=panel.smooth)
```

# Preprocessing

## Transformations

## Split the dataset

```{r}

```


# Models

```{r}

```

