---
title: "Final_Project"
author: "Nicola Cortinovis, Roberta Lamberti, Marta Lucas, Stefano Tumino"
date: "`r Sys.Date()`"
output: html_document
---
# Initial setup
```{r}
#install.packages("mgcv")
#install.packages("dplyr")
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("ipred")
#install.packages("caret")
#install.packages("stringr")
#install.packages("ggcorrplot")
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("GGally")

```

## Import libraries
```{r}
library(stringr)
library(mgcv)        # GAMs
library(rsample)     # data splitting
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)
library(ggcorrplot)
library(tidyverse)
library(ggplot2)
library(GGally)
```

## Import dataset
```{r}
df <- read.csv("phones_data.csv", header=T)

summary(df)
```
## Dataset 

| Variables      |       Description                            |    Type     |
| -------------  | -------------------------------------------  |-------------|
| brand_name     | Name of the phone brand                      |     chr     |
| model_name     | Name of the phone model                      |     chr     |
|    os          |    Operating system                          |     chr     |
| Popularity     | The popularity of the phone in range 1-1224  |     int     |
| Best Price     | Best price of the price-range in (UAH)       |     num     |
| Lowest Price   | Highest price of the price-range in (UAH)    |     num     |
| highest_price  | Lowest price of the price-range in (UAH)     |     num     |
| Sellers amount |   The amount sellers sold the phone          |     num     |
| Screen size    |  The size of phone's screen (inches).        |     num     |
| Memory size    |  The size of the phone's memory (GB)         |     num     |
| Battery size   |  The size of the phone's battery (mAh)       |     num     |
| Release date   |  The launch date of the product on the market|     chr     |

## Variable cleaning

### Remove useless variables
TODO: elenco variabili categoriali e spiegazioni sul dataset
```{r}
# rimozione colonna degli indici
df$X <- NULL
```

### Factorize the categorical variables
```{r}
df$brand_name <- factor(df$brand_name)
df$os <- factor(df$os)
```

### Convert the prices to euros
```{r}

df$best_price <- df$best_price*0.024
df$lowest_price <- df$lowest_price*0.024
df$highest_price <- df$highest_price*0.024
```

### Make the os variable more readable

```{r}
levels(df$os) <- c("other", "Android", "Android", "iOS", "other", "Android", "Android")
```

### Gestione Meizu outlier
```{r}
df[df$os == "other", ][which.max(df$memory_size[df$os == "other"]),] # soft check
df[df$os == "other", ][which.max(df$memory_size[df$os == "other"]),]$os = "Android"
```

### Split the timestamps
```{r}
df$month <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][1]}))
df$year <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][2]}))-2000
```

# Dataset cleaning 
- sistema operativo mancante modelli molto vecchi di telefoni, non sono smartphone 
- memory size mancante per gli stessi 

## Remove duplicated rows
```{r}

idxs <- which(duplicated(df[,c(-2, -4)]))

succ <- idxs[-length(idxs)] - idxs[-1] == -1
succ <- c(succ, FALSE)

i = 1
while (i <= length(idxs)){
  start = idxs[i]-1
  sum <- c(df$popularity[idxs[i]])
  while (succ[i] == TRUE){
    i = i + 1
    sum <- c(sum, df$popularity[idxs[i]])
  }
  df$popularity[start] <- mean(c(sum, df$popularity[start]))
  i = i + 1
}

df <- df[-idxs, ]

```

 
Let's divide the popularity into levels,assigning three levels associated with a numeric score. (Da mettere dopo, forse considerare split 25/50/25 (low/med/high) con i quartili)
- From 0 to 25% -> low
- From 25% to 50% -> medium-low
- From 50% to 75% -> medium-high
- From 75% to 100% -> high

```{r}
df$popularity <- as.numeric(df$popularity)

# tag <- c(0, quantile(df$popularity, 0.25), quantile(df$popularity, 0.5),quantile(df$popularity, 0.75), Inf)

tag <- quantile(df$popularity)

df$popularity_levels <- cut(df$popularity, breaks = tag,
labels=c("low", "medium-low", "medium-high", "high"), include.lowest=TRUE)

```

## Manage missing values
Since there are only 10 missing values for battery size and 2 NAs for screen size, we can drop those observations.

We notice that most of the missing values for memory size are relative to phones that are not smart-phones, for this reason we decided to substitute them with the median of the values for memory size of non-smartphones.TODO: sistemare
For the highest/lowest price we'll just substitute the mean of those columns.

### Drop Battery and Screen size missing values rows
```{r}
df <- df[- which(is.na(df$battery_size)),]
df <- df[- which(is.na(df$screen_size)),]
```

### Impute memory size missing values

We notice that most of the missing values are not smartphones... (os) TODO: spiegare quello che abbiamo fatto per la media
```{r}
tmp <- df[which(df$os == "other"),]$memory_size
df$memory_size[which(is.na(df$memory_size))] <- median(tmp[-which(is.na(tmp))])
```


### substitute the highest/lowest price NA with the median of the column
```{r}
tmp <- which(is.na(df$lowest_price))
df$lowest_price[tmp] <- median(df$lowest_price[-tmp])
```

```{r}
tmp <- which(is.na(df$highest_price))
df$highest_price[tmp] <- median(df$highest_price[-tmp])

```

### Remove useless variables
(le possiamo tenere per mantenere le info delle righe, ma non usiamo nei modelli)
```{r}
# remove useless variables
#df$model_name <- NULL
#df$release_date <- NULL
#df$popularity <- NULL
```

# Data Analysis
## Correlation matrix
```{r}
corr <- cor(df[, c(-1,-2, -3, -12, -13, -14, -15)], use="complete.obs" )

ggcorrplot(corr, hc.order = TRUE, lab = TRUE, colors=c("yellow", "white", "purple"))
```
TODO: explain what columns need to be removed (highest/lowest price)
## Single plots and Outlier analysis
### Numerical
TODO: controllare i plot nuovi
```{r}
cols <- c(5, 8:11)

for (i in cols) {
  p1 <- ggplot(df, aes_string(x=colnames(df)[i])) + geom_boxplot(fill="yellow", width= 15, position = position_nudge(y=300))+ 
  geom_histogram(fill="purple", color = "black") + ggtitle(toupper(colnames(df)[i])) + theme(plot.title = element_text(hjust = 0.5, size=16, face="bold.italic"))
  plot(p1)
}
```

### Categorical

Capire come gestire i brand names nei plot delle categoriali. Ricordarsi dell'interpretazione del best price al variare dei mesi (09/10 release Iphone , 02 release molti Samsung)

#### Provare test set 2021
TODO: controllare il codice 
```{r}

for (i in c(1,3,13,14)) {
  p1 <- ggplot(data = data.frame(x = sort(names(table(df[, i]))), y = as.numeric(table(df[, i]))),
         aes(x = x, y = y, fill = x)) +
    geom_bar(fill= "purple",stat = "identity") + ggtitle(toupper(colnames(df)[i])) + labs(x = colnames(df)[i], y = "count") + theme(plot.title = element_text(hjust = 0.5, size=16, face="bold.italic"))
  plot(p1)
}
#  NB - non si riescono a riordinare i mesi, probabilmente li sta riordinando come caratteri alfanumerici, non interi
```

```{r}
for (i in c(1, 3, 13, 14, 15)) {
    p1 <- ggplot(df, aes(x = factor(df[, i]), y = df$best_price, fill = factor(df[, i]))) +
  geom_boxplot(fill = "yellow") + labs(x = colnames(df)[i], y = "best price") + ggtitle(toupper(colnames(df)[i])) + theme(plot.title = element_text(hjust = 0.5, size=16, face="bold.italic"))
    plot(p1)
}
```

### Pairs plot
```{r fig.height=7.5, fig.width=7.5, warning=FALSE}}

df_subset <- df[cols]

ggpairs(df_subset, columns = 1:length(cols), lower = list(geom_smooth()))

 #pairs(df[cols], pch=19, lower.panel=panel.smooth) # meglio con pairs?
```

# Preprocessing

## Transformations

### Log transformation
```{r}
for (i in c(5:8, 11)){
  df[, i] <- log(df[, i])
}
```

### Log2 transformation
```{r}
df$memory_size <- floor(log2(df$memory_size*1e4))
```

### Add categorical variable

The variables memory size and screen size are bimodal, we can add a categorical variable which differentiate smartphones from phones.
We can do this setting to 1 (smartphone) if the memory size is greater than 14 and the screen size is greater than 3.5, 0 (phones) otherwise.


```{r}
df$phone_type <- factor((df$memory_size < 14), labels = c("smartphone", "phone"))
```

### Check me
```{r}
cols <- c(5, 8:11)

for (i in cols) {
  p1 <- ggplot(df, aes_string(x=colnames(df)[i])) + geom_boxplot(fill="yellow", width= 15, position = position_nudge(y=300))+ 
  geom_histogram(fill="purple", color = "black") + ggtitle(colnames(df)[i]) + theme(plot.title = element_text(hjust = 0.5))
  plot(p1)
}
```

Da vedere se ha senso separare (solo durante la data analysis) le due "famiglie"
Da fare un ciclo per vedere tutte le var al cambiare di phone type.
```{r}
cols <- c(5, 8:11, 13, 14)
for (i in cols){
  layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
  par(mar=c(0,3.1,1.1,2.1))
  boxplot(df[,i] ~ df$phone_type, horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main=colnames(df)[i], ylim=c(min(df[,i])-1, max(df[,i])+1))
  par(mar=c(4,3.1,1.1,2.1))
  hist(df[which(df$phone_type == "smartphone"), i], main = "", breaks = seq(min(df[,i]-1), max(df[,i])+1), col="purple", xlim=c(min(df[,i])-1, max(df[,i])+1))
  hist(df[which(df$phone_type == "phone"), i], breaks = seq(min(df[,i]-1), max(df[,i])+1), main = "", col="yellow", add=T)
}
```

```{r}
layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
par(mar=c(0,3.1,1.1,2.1))
boxplot(df$memory_size ~ df$phone_type, horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main="memory_size", ylim=c(4, 23))
par(mar=c(4,3.1,1.1,2.1))
hist(df$memory_size[which(df$phone_type == "smartphone")], breaks=5:25, main = "", col="purple", xlim=c(4, 23))
hist(df$memory_size[which(df$phone_type == "phone")], breaks=5:25, main = "", col="yellow", add=T)
```

```{r}
layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
par(mar=c(0,3.1,1.1,2.1))
boxplot(df$battery_size ~ df$phone_type, horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main="battery_size", ylim=c(5.5, 10.5))
par(mar=c(4,3.1,1.1,2.1))
hist(df$battery_size[which(df$phone_type == "smartphone")], breaks=seq(from=5, to=10, by=0.25), main = "", col="purple", xlim=c(5.5, 10.5))
hist(df$battery_size[which(df$phone_type == "phone")], breaks=seq(from=5, to=10, by=0.25), main = "", col="yellow", add=T)
```
```{r}
layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
par(mar=c(0,3.1,1.1,2.1))
boxplot(df$screen_size ~ df$phone_type, horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main="screen_size", ylim=c(0, 12))
par(mar=c(4,3.1,1.1,2.1))
hist(df$screen_size[which(df$phone_type == "smartphone")], breaks=seq(from=0, to=10, by=1), main = "", col="purple", xlim=c(0, 12))
hist(df$screen_size[which(df$phone_type == "phone")], breaks=seq(from=0, to=10, by=1), main = "", col="yellow", add=T)
```

```{r}
tmp <- sort(table(df$brand_name))

barplot(tmp, main="brand_names", col="purple")
#abline(h=40, col="red", lwd=2)
#abline(h=20, col="blue", lwd=2)
abline(h=40, col="pink", lwd=4)
#abline(h=10, col="green", lwd=2)
#abline(h=1, col="yellow", lwd=2)
#abline(h=5, col="orange", lwd=2)


```
```{r}
c("Total"=sum(tmp),
"Leq 5"=sum(tmp[tmp <= 40]),
"Proportion (%)"=sum(tmp[tmp <= 40])/sum(tmp) * 100)


```

#### Other
```{r}
to_remove <- names(tmp[tmp <= 15])

vals <- c()
for (i in 1:length(levels(df$brand_name))){
  if (any(levels(df$brand_name)[i] == to_remove)){
    vals <- c(vals, "other")
  } else {
    vals <- c(vals, levels(df$brand_name)[i])
  }
}

levels(df$brand_name) <- vals
table(vals)
```

#### Drop
```{r}
#to_remove <- names(tmp[tmp <= 5])

#idxs <- c()
#for (i in 1:length(df$brand_name)){
#  if (any(df$brand_name[i] == to_remove)){
#    idxs <- c(idxs, i)
#  }
#}

#df <- df[-idxs,]

#vals <- c()
#for (i in 1:length(levels(df$brand_name))){
#  if (any(levels(df$brand_name)[i] == to_remove)){
#    vals <- c(vals, "Samsung")
#  } else {
#    vals <- c(vals, levels(df$brand_name)[i])
#  }
#}

#¯levels(df$brand_name) <- vals
```

```{r}
tmp <- sort(table(df$brand_name))

barplot(tmp, main="brand_names", col="purple")
```


## Train-Test Split
Come gestiamo le info temp? Ha senso standardizzarle? Forse campioniamo. 
```{r}
split <- initial_split(df, prop = 0.7)
train <- training(split)
test <- testing(split)
```

```{r}
df$epoch <- (df$year - 13)*12 + df$month

sorted_df <- df[sort(df$epoch, index.return=T)$ix,]

train <- sorted_df[1:round(nrow(sorted_df) * 0.8),]
test <- sorted_df[round(nrow(sorted_df) * 0.8):nrow(sorted_df),]

for (i in 8:11){
  mu  <- mean(train[, i])
  std <- sd(train[, i])
  
  train[, i] <- (train[, i] - mu) / std
  test[, i] <- (test[, i] - mu) / std
}
```

# Model fitting

```{r}
summary(df)
summary(test)
```



## GAM
### Default
```{r fig.height=5, fig.width=5}
# all variables without splines
fit_gam0 <- gam(best_price ~ os + sellers_amount+ phone_type + screen_size + memory_size + battery_size + month + year + popularity_levels + brand_name , data = train)
summary(fit_gam0)

y_hat <- predict(fit_gam0, newdata = train)
pred <- predict(fit_gam0, newdata = test)
results0<- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam0))
```
```{r}
# without month, year, sellers amount, popularity levels, phone type
fit_gam <- gam(best_price ~ os + screen_size + memory_size + battery_size + brand_name, data = train)
summary(fit_gam)

y_hat <- predict(fit_gam, newdata = train)
pred <- predict(fit_gam, newdata = test)
results_0<- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam))
```

```{r fig.height=5, fig.width=5}
# add month to the previous model
fit_gam1 <- gam(best_price ~ os + screen_size + memory_size + battery_size + month +  brand_name, data = train)
summary(fit_gam1)

y_hat <- predict(fit_gam1, newdata = train)
pred <- predict(fit_gam1, newdata = test)
results1<- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam1))
```

```{r}
# funzione non vista a lezione
par(mfrow=c(2,2))
gam.check(fit_gam1)
```

### With Splines
```{r}
# all continuos variables with splines, also all categorical variables included
fit_gam2 <- gam(best_price ~ os + s(screen_size) + s(memory_size) + s(battery_size) + s(sellers_amount) + s(month)+  phone_type + popularity_levels + brand_name, data = train, method = "REML")
summary(fit_gam2)

y_hat <- predict(fit_gam2, newdata = train)
pred <- predict(fit_gam2, newdata = test)
results2 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam2))
```
Smooth effects
```{r}
par(mfrow=c(2,3))
plot(fit_gam2)
```
```{r}
par(mfrow=c(2,3))
plot(fit_gam2, residuals = TRUE, pch = 19, cex = 0.8)
```
```{r}
par(mfrow=c(2,2))
gam.check(fit_gam2) # funzione nn vista a lezione, credo

```

```{r}
# remove brand name, to see what happens
fit_gam3 <- gam(best_price ~ os + s(screen_size) + s(memory_size) + s(battery_size) + month + phone_type + popularity_levels, data = train, method = "REML")
summary(fit_gam3)

y_hat <- predict(fit_gam3, newdata = train)
pred <- predict(fit_gam3, newdata = test)
results3 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam3))
```
### model with splines, without phone type, month, popularity levels, sellers amount
```{r}
fit_gam4 <- gam(best_price ~ os + s(screen_size) + s(memory_size) + s(battery_size) + brand_name, data = train, method = "REML")
summary(fit_gam4)

y_hat <- predict(fit_gam4, newdata = train)
pred <- predict(fit_gam4, newdata = test)
results4 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam4))
```
### With selction of splines and just brand_name
```{r}
# with splines and brand_name
fit_gam5 <- gam(best_price ~ s(screen_size) + s(memory_size) + s(battery_size) + brand_name, data = train, method = "REML")
summary(fit_gam5)

y_hat <- predict(fit_gam5, newdata = train)
pred <- predict(fit_gam5, newdata = test)
results5 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam5))
```
```{r}
par(mfrow=c(1,3))
plot(fit_gam5)
```

```{r}
par(mfrow=c(1,3))
plot(fit_gam5, residuals = TRUE, pch = 19, cex = 0.8)
```
```{r}
par(mfrow=c(2,2))
gam.check(fit_gam5)
```

```{r}
# interactions with os
fit_gam6 <- gam(best_price ~  s(screen_size, by = os) + s(memory_size, by= os) + s(battery_size, by= os) + brand_name + os, data = train, method = "REML")
summary(fit_gam6)

y_hat <- predict(fit_gam6, newdata = train)
pred <- predict(fit_gam6, newdata = test)

results6 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam6))
```

```{r}
# interactions with phone type
fit_gam7 <- gam(best_price ~  s(screen_size, by = phone_type) + s(memory_size, by= phone_type) + s(battery_size, by= phone_type) + brand_name + phone_type, data = train, method = "REML")
summary(fit_gam7)

y_hat <- predict(fit_gam7, newdata = train)
pred <- predict(fit_gam7, newdata = test)
results7 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam7))
```

```{r}
# interactions with popularity for sellers amount,and month, while with os for the others
fit_gam8 <- gam(best_price ~  
                  s(screen_size, by = os) + s(memory_size, by= os) + s(battery_size, by= os) + 
                  s(sellers_amount, by = popularity_levels) + s(month, by = popularity_levels) + 
                  brand_name + os + popularity_levels, data = train, method = "REML")
summary(fit_gam8)

y_hat <- predict(fit_gam8, newdata = train)
pred <- predict(fit_gam8, newdata = test)
results8 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam8))
```

```{r}
par(mfrow=c(2,2))
gam.check(fit_gam8)
```

```{r}
# interactions between continuous variables
fit_gam9 <- gam(best_price ~  s(screen_size, memory_size) + s(battery_size, memory_size) + brand_name , data = train, method = "REML")

summary(fit_gam9)

y_hat <- predict(fit_gam9, newdata = train)
pred <- predict(fit_gam9, newdata = test)
results9 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam9))
```

```{r}
# interactions between continuos variables and continuos with categorical
fit_gam10 <- gam(best_price ~  s(screen_size, memory_size) + s(battery_size, memory_size)+ 
                 s(screen_size, by = os)  + s(memory_size, by = os) +s(battery_size, by= os) + 
                  s(sellers_amount, by = popularity_levels) + s(month, by = popularity_levels) + 
                  brand_name + os + popularity_levels, data = train, method = "REML")

summary(fit_gam10)

y_hat <- predict(fit_gam10, newdata = train)

pred <- predict(fit_gam10, newdata = test)
results10 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam10))
```

```{r}
# concatenate results in a matrix
results <- rbind(results0,results_0,results1, results2, results3, results4, results5, results6, results7, results8, results9, results10)
rownames(results) <- c("GAM0","GAM_0","GAM1", "GAM2", "GAM3", "GAM4", "GAM5", "GAM6" , "GAM7", "GAM8", "GAM9", "GAM10")
colnames <- c("RMSE", "R2", "RMSE_test", "R2_test", "AIC")
results
```
The model has to take the brand_name into account, it seems to be the most important categorical variable.
If we don't take it into account (GAM3) we lose a lot of information.
The best models without interactions are GAM5 and GAM4, with the lowest RMSE and AIC.Perhaps the best is indeed GAM5, since it has the least covariates and the highest R2.

Interactions seem to be important! GAM 6,7,8 The best result is GAM8, we considered the interactions
between month,sellers amount with popularity levels ans screen size, memory size, battery size with os.
Note: -la standardizzazione del train set va been? penso di si, anche se ci sono q.tà negative?
- il dato temporale sembra non essere particolarmente importante in questi modelli, c'è qualcosa di 
sbagliato? -> forse si può effettivamente provare a fare un trainset random 
- Non ho fatto il grafico dei residui di tutto se no , buonanotte 

- Custom ggplots, cambiare lo sfondo etc.. x plot + carini (da proporre a Nicola)











