---
title: "Final_Project"
author: "Nicola Cortinovis, Roberta Lamberti, Marta Lucas, Stefano Tumino"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r, include = FALSE}
#install.packages("mgcv")
#install.packages("dplyr")
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("ipred")
#install.packages("caret")
#install.packages("stringr")
#install.packages("ggcorrplot")
#install.packages("cowplot")
```

```{r, include = FALSE}
library(stringr)
library(mgcv)        # GAMs
library(rsample)     # data splitting
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)
library(ggcorrplot)
library(ICglm)
library(randomForest)
library(earth)
library(gridExtra)
library(cowplot)
```


# Introduction

The aim of this project is to predict the price of a smartphone based on its features. The dataset used for this analysis is the [mobile phones dataset](https://www.kaggle.com/datasets/artempozdniakov/ukrainian-market-mobile-phones-data), which contains 13 columns and 1224 rows:

| Variables | Description                                   | Type          |
| -------------  | -------------------------------------------   | ------------- |
| X              | Index of the phone                            |     int       |
| Brand_Name     | Name of the phone brand                       |     chr       |
| Model_Name     | Name of the phone model                       |     chr       |
|    Os          |    Operating system                           |     chr       |
| Popularity     | The popularity of the phone in range 1-1224   |     int       |
| Best_Price     | Best price of the price-range in (UAH)        |     num       |
| Lowest_Price   | Highest price of the price-range in (UAH)     |     num       |
| Highest_Price  | Lowest price of the price-range in (UAH)      |     num       |
| Sellers_Amount |   The amount sellers sold the phone           |     num       |
| Screen_Size    |  The size of phone's screen (inches).         |     num       |
| Memory_Size    |  The size of the phone's memory (GB)          |     num       |
| Battery_Size   |  The size of the phone's battery (mAh)        |     num       |
| Release_Date   |  The launch date of the product on the market |     chr       |


```{r Intoduction}
df <- read.csv("phones_data.csv", header=T)
head(df)
```

Specifically, our objective is to predict the `best_price` variable. Our approach consists of the following steps:

# Data exploration

## Data preprocessing

Firstly we remove from the dataset the index column `X`
```{r}
df$X <- NULL
```

Our next step is to briefly explore the `chr` variables and transform the appropriate ones into factors.

 - `model_name` won't be transformed into a factor because it has too many levels, almost a unique model_name for each row;
 
 - `brand_name` will be transformed into a factor;
 
 - `os`  will be transformed into a factor;
 
 - `release_date`  won't be transformed into a factor because it will be used to create two new variables: `month` and `year`. 
 
```{r}
length(unique(df$model_name))
```

```{r}
df$brand_name <- factor(df$brand_name)
df$os <- factor(df$os)
```

```{r}
df$month <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][1]}))
df$year <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][2]}))-2000
```

For clarity's sake we convert the ukrainian currency (UAH) into euros (€) and rename the blank "" `os` class as "other".

```{r}
df$best_price <- df$best_price*0.024
df$lowest_price <- df$lowest_price*0.024
df$highest_price <- df$highest_price*0.024

levels(df$os)[1] <- "other"
```

We decide to investigate the `os` variable

```{r}
table(df$os)
```
Given the insufficient amount of data for the `EMUI`, `KAIOS`, `OxygenOS` and `WindowsPhone` factor levels, we decide to aggregate them into the `other` and `Android` levels based on their characteristics.
```{r}
levels(df$os) <- c("other", "Android", "Android", "iOS", "other", "Android", "Android")
```

```{r}
summary(df)
```

From the summary we notice the presence of some NAs in the `battery_size`, `screen_size`, `memory_size`, `lowest_price` and `highest_price` variables. Given the small amount of NAs in the first two we  decide to remove the rows with NAs. Further investigation into the `memory_size` shows that the NAs are present only for the "other" `os` class, so we decide to fill them with the median of the `memory_size` for the "other" `os` class. The choice of the median is dictated by the presence of a few outliers.

```{r battery_size and screen_size NAs}
df <- df[- which(is.na(df$battery_size)),]
df <- df[- which(is.na(df$screen_size)),]
```

```{r memory_size NAs}
tmp <- df[which(df$os == "other"),]$memory_size
df$memory_size[which(is.na(df$memory_size))] <- median(tmp[-which(is.na(tmp))])


```

The variables `lowest_price` and `highest_price` also show some outliers so their NAs have been filled with their median.

```{r lowest_price and highest_price NAs}
tmp <- which(is.na(df$lowest_price))
df$lowest_price[tmp] <- median(df$lowest_price[-tmp])
tmp <- which(is.na(df$highest_price))
df$highest_price[tmp] <- median(df$highest_price[-tmp])

summary(df)
```

The dataset contains several duplicates, where phones share the same characteristics but have different popularity levels. Therefore, we eliminate these duplicate observations as they provide redundant information. This process involves retaining only the first occurrence of each duplicate and replacing its popularity with the average popularity of the duplicates. An example is given below.

```{r duplicates motivation}
df[2:4,]
```
```{r duplicates handling}
# Find the indices of duplicate rows
idxs <- which(duplicated(df[,-c(2, 4)]))

# Check if each index is succeeded by the next one in the sequence
succ <- c(idxs[-1] - idxs[-length(idxs)] == 1, FALSE)

i = 1
while (i <= length(idxs)){
  start = idxs[i]
  sum <- c(df$popularity[start])
  while (succ[i]){
    i = i + 1
    sum <- c(sum, df$popularity[idxs[i]])
  }
  df$popularity[start] <- mean(sum)
  i = i + 1
}

# Remove the duplicate rows
df <- df[-idxs, ]

# Remove the model_name column
df$model_name <- NULL
```

```{r Meizu outlier, include=FALSE}
df[df$os == "other", ][which.max(df$memory_size[df$os == "other"]),]$os = "Android"
```

The `popularity` variable is unique for each row, therefore we decided to create a new variable `popularity_levels` which divides the popularity into 4 classes: "low", "medium", "high" and "very high" based on the quartiles of the `popularity` variable.

```{r}
df$popularity <- as.numeric(df$popularity)

tag <- quantile(df$popularity)

df$popularity_levels <- cut(df$popularity, breaks = tag,
labels=c("low", "medium", "high", "very high"), include.lowest=TRUE)

df$popularity <- NULL
```

## Data visualization

```{r}
corr <- cor(df[, c("battery_size", "memory_size", "screen_size", "best_price", "highest_price", "lowest_price", "sellers_amount")], use="complete.obs" )

ggcorrplot(corr, hc.order = TRUE, lab = TRUE, colors = c("#AFDDD5", "#EFDEC0", "#FF284B"))
```

From the correlation plot, we observe a strong positive linear correlation between best_price, highest and lowest price. For this reason we don't consider them in our analysis.

```{r}
df$highest_price <- NULL
df$lowest_price <- NULL
```

```{r, message=FALSE, warning=FALSE}
cols <- c("battery_size", "memory_size", "screen_size", "sellers_amount")

plots = list()

for (i in cols) {
  p1 <- ggplot(df, aes_string(x=i)) + geom_boxplot(fill="yellow", width= 15, position = position_nudge(y=300))+ 
    geom_histogram(fill="purple", color = "black") + ggtitle(toupper(i)) + theme(plot.title = element_text(hjust = 0.5, size=12, face="bold.italic"))
    plots <- c(plots, list(p1))}

grid.arrange(grobs=plots, ncol=2, nrow=2)
```
The `battery_size`, `memory_size` and `sellers_amount` covariates plots all show the presence of right-skewed distributions. For those we decided to apply a logarithmic transformation, notably for `memory_size` we choose a logarithm of base 2. The `screen_size` covariate too displays a certain degree of right-skewness, but also a clear bi-modal distribution. For 
this reason we decided to leave it as it is.

```{r}
df$log_battery_size <- log(df$battery_size)
df$log_memory_size <- floor(log2(df$memory_size*1e4))
df$log_sellers_amount <- log(df$sellers_amount)

```

```{r, message=FALSE, warning=FALSE}
cols <- c("log_battery_size", "log_memory_size", "log_sellers_amount")

plots = list()

for (i in cols) {
  p1 <- ggplot(df, aes_string(x=i)) + geom_boxplot(fill="yellow", width= 15, position = position_nudge(y=550))+ 
    geom_histogram(fill="purple", color = "black", bins = 12) + ggtitle(toupper(i)) + theme(plot.title = element_text(hjust = 0.5, size= 12, face="bold.italic"))
    plots <- c(plots, list(p1))}

grid.arrange(grobs=plots, ncol=3, nrow=1)
```
The transformations seem to handle the right-skewness of the covariates. The `log_battery_size` covariate is now approximately normally distributed. The `log_memory_size` covariate exhibits a bimodal behavior, but the transformation has reduced the right-skewness. The `log_sellers_amount` is now approximately uniform. 

We further investigate the bimodal behavior by plotting the three variables we're interested in by factoring the `os` type as well.

```{r}
bimodal_vars <- c("log_battery_size", "log_memory_size", "screen_size")
df_long <- reshape2::melt(df, id.vars = "os", measure.vars = bimodal_vars)

colors <- c("other" = "deepskyblue4", "Android" = "darkorange", "iOS" = "skyblue") 

# Create the histograms
p <- ggplot(df_long, aes(x = value, fill = os)) +
  geom_histogram(color = "black", position = "identity", bins = 12) +
  scale_fill_manual(values = colors) +
  facet_wrap(~variable, scales = "free") +
  theme(legend.position = "bottom", strip.text = element_text(size = 12, face = "bold.italic"))

# Print the plot
print(p)
```

From these plots we see that there exist a clear difference in these variables distribution that's based on the type of operating system (`os`). `Android` and `iOs` categories refer to smartphones with more modern features, while `other` category refers to old-fashioned mobile phones. 

## Categorical variables exploration

```{r brand_name}
tmp <- sort(table(df$brand_name))

# Convert the table to a data frame for plotting
tmp_df <- data.frame(Brand = names(tmp), Count = as.vector(tmp))

# Create a bar plot
p <- ggplot(tmp_df, aes(x = reorder(Brand, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Brand") +
  ylab("Count") +
  ggtitle("Brand Counts") + geom_hline(yintercept = 60, color = "red", linetype = "dashed") + theme(plot.title = element_text(hjust = 0.5, size=12, face="bold.italic"))

# Print the plot
print(p)
```
As there are too many classes with an exiguous amount of observations, we decided to group the brands into 4 categories: `Samsung`, `Xiaomi`, `Apple` and `Remaining`.

```{r}
tmp <- sort(table(df$brand_name))
cut_line <- 60
to_remove <- names(tmp[tmp <= cut_line])

vals <- c()
for (i in 1:length(levels(df$brand_name))){
  if (any(levels(df$brand_name)[i] == to_remove)){
    vals <- c(vals, "Remaining")
  } else {
    vals <- c(vals, levels(df$brand_name)[i])
  }
}

levels(df$brand_name) <- vals

table(vals)
```

```{r}
tmp <- sort(table(df$brand_name))

# Convert the table to a data frame for plotting
tmp_df <- data.frame(Brand = names(tmp), Count = as.vector(tmp))

# Create a bar plot
p <- ggplot(tmp_df, aes(x = reorder(Brand, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  theme(axis.text.x = element_text(hjust = 0.5)) +
  xlab("Brand") +
  ylab("Count") +
  ggtitle("Brand Counts") + theme(plot.title = element_text(hjust = 0.5, size=12, face="bold.italic"))

# Print the plot
print(p)
```


```{r}
plots = list()
for (i in c("month","year")) {
  p1 <- ggplot(data = data.frame(x = factor(sort(as.numeric(names(table(df[, i]))))), y = as.numeric(table(df[, i]))),
         aes(x = x, y = y, fill = x)) +
    geom_bar(fill= "steelblue",color = "black", stat = "identity") + ggtitle(toupper(i)) + labs(x = i, y = "count") + theme(plot.title = element_text(hjust = 0.5, size=12, face="bold.italic"))
  plots <- c(plots, list(p1))}
grid.arrange(grobs=plots, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")
```

In analyzing the `month` plot, we observe a notable increase in the number of phones sold during the months of September (9) and October (10). 
This surge aligns with the annual release of iPhones, suggesting a pattern linked to this event. 

Additionally, the exponential growth trend evident in the `year` plot likely reflects the increasing demand for smartphones on a yearly basis. 
The peak observed in 2020 may be attributed to the COVID-19 pandemic, which significantly heightened the need for remote communication, during isolation for each member within most households. 
The count for 2021 appears lower since our dataset registers data until February.

```{r}
tmp <- sort(table(df$os))

# Convert the table to a data frame for plotting
tmp_df <- data.frame(Os = names(tmp), Count = as.vector(tmp))

# Create a bar plot
p <- ggplot(tmp_df, aes(x = reorder(Os, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  theme(axis.text.x = element_text(hjust = 0.5)) +
  xlab("Os") +
  ylab("Count") +
  ggtitle("Os counts") + theme(plot.title = element_text(hjust = 0.5, size=12, face="bold.italic"))

# Print the plot
print(p)
```
It's evident that the `Android` operating system is the most chosen in the market.

```{r}

cat_variables <- c("os", "month", "year", "popularity_levels")

plots = list()



for (i in cat_variables) {
  p1 <- ggplot(df, aes_string(x= factor(df[,i]), y = df$best_price)) + geom_boxplot(fill="indianred1") + labs(x = i, y = "best_price") + theme(plot.title = element_text(hjust = 0.5, size= 12, face="bold.italic")) + ggtitle(toupper(i))
    plots <- c(plots, list(p1))}

grid.arrange(grobs=plots, ncol=2, nrow=2)

```

From the boxplots we observe that:

  - For the `os` variable the `iOS` operating system shows a significantly higher `best_price` than the others. We also notice that the `Android` level has several outliers;
  - For the `month` variable, the `best_price` is higher in the months of September and October, which aligns with the release of new iPhones. In the month of February, the `best_price` is also higher than the others because of the release of the flagships Samsung smartphones;
  - For the `year` variable, the `best_price` exhibits a growing trend over the years;
  - For the `popularity_levels` variable, the `best_price` grows with the popularity, except for some outliers.
  
## Training and testing datasets split 

```{r}
# df$epoch <- (df$year - 13) * 12 + df$month
# 
# sorted_df <- df[sort(df$epoch, index.return=T)$ix,]
# 
# train <- sorted_df[1:round(nrow(sorted_df) * 0.8),]
# test <- sorted_df[round(nrow(sorted_df) * 0.8):nrow(sorted_df),]

set.seed(15)

split <- initial_split(df, prop = 0.8)
train <- training(split)
test <- testing(split)
```

We decide to split the dataset into a training set, which will be used to train the model, and a testing set, which will be used to evaluate the model's performance. The split is done in a chronological order, with 80% of the data being used for training (`01/2013 - 09/2020`) and the remaining 20% for testing (`09/2020 - 02/2021`). This is done with the idea of ensuring that the model is trained on data that precedes the testing data, which is more realistic and might help the model to generalize better. Our empirical trials have shown a slight improvement over the random split.

# Linear models

We begin our analysis by fitting a linear model to the data. We test our linear models by computing the following statistics:

  - RMSE on the training set;
  - R2 on the training set;
  - RMSE on the test set;
  - R2 on the test set;
  - AIC;
  - BIC;
  - Training time.

We follow a top-down approach to select the most significant variables.

### Model 1

Our first model includes all the variables with the applied transformations.

```{r warning=FALSE}

start <- Sys.time()
lm_model_1 <- lm(best_price ~ brand_name + os + log_sellers_amount + screen_size + log_memory_size + log_battery_size + month + year + popularity_levels, data = train)

end <- Sys.time()

summary(lm_model_1)

y_hat <- predict(lm_model_1, newdata = train)
pred <- predict(lm_model_1, newdata = test)

result1 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(lm_model_1), "BIC" = BIC(lm_model_1), "Time"=end-start)
result1
```

The results show that the variables: `month`, `popularity_levels`, `log_sellers_amount` and `log_battery_size` don't affect the model as much. Our next models will exclude these variables.

### Model 2

```{r}
start <- Sys.time()
lm_model_2 <- lm(best_price ~ brand_name + os + screen_size + log_memory_size + year, data = train)

end <- Sys.time()

summary(lm_model_2)

y_hat <- predict(lm_model_2, newdata = train)
pred <- predict(lm_model_2, newdata = test)

result2 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(lm_model_2), "BIC" = BIC(lm_model_2), "Time"=end-start)
result2
```

The results show that the model's performance is slightly worsened by excluding the variables `log_sellers_amount`, `log_battery_size` and `month`. The next model will include interactions between some variables to check if it's possible to improve the model's performance.

### Model 3

```{r}
start <- Sys.time()

lm_model_3 <- lm(best_price ~ brand_name + os + screen_size + year + os:log_memory_size  + os:screen_size, data = train)


end <- Sys.time()

summary(lm_model_3)

y_hat <- predict(lm_model_3, newdata = train)
pred <- predict(lm_model_3, newdata = test)

result3 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(lm_model_3), "BIC" = BIC(lm_model_3),  "Time"=end-start)
result3
```

The results show that the model's performance is improved by including the interactions between the variables. However there's still a non-significant variable, namely `screen_size`. The next model will therefore exclude the `screen_size` variable.

### Model 4

```{r}
start <- Sys.time()

lm_model_4 <- lm(best_price ~ brand_name + os + year + os:log_memory_size  + os:screen_size, data = train)

end <- Sys.time()

summary(lm_model_4)

y_hat <- predict(lm_model_4, newdata = train)
pred <- predict(lm_model_4, newdata = test)

result4 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(lm_model_4), "BIC" = BIC(lm_model_4), "Time"=end-start)
result4
```


The results for the final model are unchanged compared to the previous one. 

### Residual analysis for model 4

```{r}
par(mfrow=c(2,2))
plot(lm_model_4)
```

From the residuals of the model we can see that there is heteroschedasticity. This means that the variance of the residuals is not constant. This is a violation of the assumption of homoscedasticity.
We can also see that the residuals are not normally distributed. Knowing that all the dependent variables are positive, we will try to fix these issues by applying a logarithmic transformation to the response variable.

## Linear models with logarithmic response variable

```{r}
df$log_best_price <- log(df$best_price)
train$log_best_price <- log(train$best_price)
test$log_best_price <- log(test$best_price)
```

### Model 5

```{r}
start <- Sys.time()

lm_model_5 <- lm(log_best_price ~ brand_name + os + month + year + os:log_memory_size + os:log_battery_size, data = train)

end <- Sys.time()

summary(lm_model_5)

y_hat <- predict(lm_model_5, newdata = train)
pred <- predict(lm_model_5, newdata = test)

result_log_5 <- c("RMSE"=RMSE(exp(y_hat), train$best_price), "R2"=R2(exp(y_hat), train$best_price), "RMSE_test"=RMSE(exp(pred), test$best_price), "R2_test"=R2(exp(pred), test$best_price), "AIC"=AIC(lm_model_5), "BIC" = BIC(lm_model_5),  "Time"=end-start)
result_log_5
```

### Model 6

```{r}
start <- Sys.time()

lm_model_6 <- lm(log_best_price ~ brand_name + os + month + os:log_memory_size + os:log_battery_size, data = train)

end <- Sys.time()

summary(lm_model_6)

y_hat <- predict(lm_model_6, newdata = train)
pred <- predict(lm_model_6, newdata = test)

result_log_6 <- c("RMSE"=RMSE(exp(y_hat), train$best_price), "R2"=R2(exp(y_hat), train$best_price), "RMSE_test"=RMSE(exp(pred), test$best_price), "R2_test"=R2(exp(pred), test$best_price), "AIC"=AIC(lm_model_6), "BIC" = BIC(lm_model_6),  "Time"=end-start)
result_log_6
```

```{r}
par(mfrow=c(2,2))
plot(lm_model_6)
```

The logarithmic model exibits less heteroschedasticity, the residuals are more normally distributed and they are also less right-skewed. The R2 and RMSE values are improved compared to the linear model.

## Results comparison

```{r}
results_lm <- rbind(result1, result2, result3, result4, result_log_5, result_log_6)
rownames(results_lm) <- c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5", "Model 6")
round(results_lm, 3)
```

# Non-linear models

## Polynomial regression
```{r}
start <- Sys.time()
poly_1 <- lm(best_price ~ brand_name + os + poly(screen_size,3) + poly(log_memory_size, 3) + poly(log_battery_size, 3), data = train)

end <- Sys.time()

summary(poly_1)

y_hat <- predict(poly_1, newdata = train)
pred <- predict(poly_1, newdata = test)

result_poly_1 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(poly_1), "BIC" = BIC(poly_1), "Time"=end-start)
result_poly_1
```

```{r}
start <- Sys.time()
poly_2 <- lm(best_price ~ brand_name + I(screen_size^2) + I(screen_size^3) + I(log_memory_size^2) + I(log_memory_size^3) + I(log_battery_size^2) + log_battery_size, data = train)

end <- Sys.time()

summary(poly_2)

y_hat <- predict(poly_2, newdata = train)
pred <- predict(poly_2, newdata = test)

result_poly_2 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(poly_2), "BIC" = BIC(poly_2), "Time"=end-start)
result_poly_2
```


## Generalized Additive Models (GAM)
## Without splines
### Model with all variables
```{r}
start <- Sys.time()
fit_gam1 <- gam(best_price ~ os + sellers_amount + screen_size + memory_size + battery_size + month + year + popularity_levels + brand_name , data = train)
end <- Sys.time()
summary(fit_gam1)


y_hat <- predict(fit_gam1, newdata = train)
pred <- predict(fit_gam1, newdata = test)
results1<- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam1), "BIC"= BIC(fit_gam1), "Time"=end-start)
```

### sellers amount, month, battery size

```{r}
start <- Sys.time()
fit_gam2 <- gam(best_price ~ os +  screen_size + memory_size + year + popularity_levels + brand_name  , data = train)
end <- Sys.time()
summary(fit_gam2)


y_hat <- predict(fit_gam2, newdata = train)
pred <- predict(fit_gam2, newdata = test)
results2<- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam2), "BIC"= BIC(fit_gam2),  "Time"=end-start)
```
### With splines
```{r}

start <- Sys.time()
fit_gam3 <- gam(best_price ~ os + s(screen_size) + s(memory_size) + s(battery_size) + s(sellers_amount) + s(month)+ year+ popularity_levels + brand_name, data = train, method = "REML")
end <- Sys.time()
summary(fit_gam3)

y_hat <- predict(fit_gam3, newdata = train)
pred <- predict(fit_gam3, newdata = test)
results3 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam3), "BIC"= BIC(fit_gam3),  "Time"=end-start)
```
We can remove the non significant variables from the model, in this case `os`. Regarding the splines we can see that the `month` variable has a linear effect on the model, so we can remove the spline term from it.

TODO: controllare se ha senso la giustificazione
```{r}
start <- Sys.time()
fit_gam4 <-gam(best_price ~  s(screen_size) + s(memory_size) + s(battery_size) + s(sellers_amount) + year + brand_name, data = train, method = "REML")
end <- Sys.time()
summary(fit_gam4)

y_hat <- predict(fit_gam4, newdata = train)
pred <- predict(fit_gam4, newdata = test)

results4 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam4), "BIC"= BIC(fit_gam4), "Time"=end-start)
```

### Smoothing effects plots
```{r}
par(mfrow=c(2,2))
plot(fit_gam4)
```
The smooth terms for `screen_size`,`battery_size`, `memory_size` and `sellers_amount` show a clear non-linear relationship with the `best_price` variable, as confirmed by the edf values in the summary of the model.

### Splines and interactions
Taking into consideration the important interactions, previously discovered in the linear models, we decided to include them in the GAM model. 
```{r}
start <- Sys.time()
fit_gam5 <- gam(best_price ~  brand_name+ year+ s(screen_size, by = os)+ s(memory_size)  +s(sellers_amount) , data = train, method = "REML")
end <- Sys.time()
summary(fit_gam5)

y_hat <- predict(fit_gam5, newdata = train)
pred <- predict(fit_gam5, newdata = test)
results5 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=AIC(fit_gam5), "BIC"= BIC(fit_gam5),  "Time"=end-start)
```
```{r}
par(mfrow=c(1,3))
plot(fit_gam5)

```

```{r}
gam_results <- rbind(results1, results2, results3, results4, results5, result_poly_1, result_poly_2)
rownames(gam_results) <- c("GAM 1", "GAM 2", "GAM 3", "GAM 4", "GAM 5", "Poly 1", "Poly 2")
round(gam_results, 3)
```


## Trees

### Regression Tree
```{r}
start <- Sys.time()
tree_1 <- rpart(best_price ~ os + sellers_amount + screen_size + memory_size + battery_size + month + year + popularity_levels + brand_name , data = train)
end <- Sys.time()
rpart.plot(tree_1)

y_hat <- predict(tree_1, newdata = train)
pred <- predict(tree_1, newdata = test)
results_tree1 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price), "AIC"=NA, "BIC"= NA, "Time"=end-start)
results_tree1
```

### Random Forest
```{r}
start <- Sys.time()
rf_1 <- randomForest(best_price ~ os + sellers_amount + screen_size + memory_size + battery_size + month + year + popularity_levels + brand_name , data = train, importance=T, proximitry=T, mtry=5)
end <- Sys.time()

y_hat <- predict(rf_1, newdata = train)
pred <- predict(rf_1, newdata = test)
results_rf1 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price),"AIC"=NA, "BIC"= NA,  "Time"=end-start)
results_rf1

varImpPlot(rf_1, sort=T, n.var=8, main="Variable Importance")
```

### Random Forest with pre-pruning
```{r}
start <- Sys.time()
rf_2 <- randomForest(best_price ~ os + sellers_amount + screen_size + memory_size + battery_size + month + year + popularity_levels + brand_name , data = train, importance=T, proximitry=T, mtry=5, maxnodes=52)
end <- Sys.time()

y_hat <- predict(rf_2, newdata = train)
pred <- predict(rf_2, newdata = test)
results_rf2 <- c("RMSE"=RMSE(y_hat, train$best_price), "R2"=R2(y_hat, train$best_price), "RMSE_test"=RMSE(pred, test$best_price), "R2_test"=R2(pred, test$best_price),"AIC"=NA, "BIC"= NA,  "Time"=end-start)
results_rf2

varImpPlot(rf_2, sort=T, n.var=8, main="Variable Importance")
```

```{r}
non_linear_results <- rbind(results1, results2, results3, results4, results5, result_poly_1, result_poly_2, results_tree1, results_rf1, results_rf2)
rownames(non_linear_results) <- c("GAM 1", "GAM 2", "GAM 3", "GAM 4", "GAM 5", "Poly 1", "Poly 2", "Tree 1", "RF 1", "RF 2")
round(non_linear_results, 3)
```








