# Quarto version

---
title: "Final_Project"
author: "Nicola Cortinovis, Roberta Lamberti, Marta Lucas, Stefano Tumino"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, include = FALSE}
# List of packages
packages <- c("stringr", "mgcv", "rsample", "dplyr", "rpart", "rpart.plot", "ipred", "caret", "ggcorrplot", "gbm", "GGally", "gridExtra")

# Function to check and install packages
check_and_install <- function(pkg){
  if(!require(pkg, character.only = TRUE)){
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

# Apply the function to the list of packages
sapply(packages, check_and_install)
```

## Import the libraries

```{r}
library(stringr)
library(mgcv)        # GAMs
library(rsample)     # data splitting
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret) 
library(ggcorrplot)
library(gbm)         # gradient boosting
library(gridExtra)
```


# Data analysis

## Dataset loading

```{r}
df <- read.csv("phones_data.csv", header=T)

summary(df)
```


| Variables      |       Description                            |    Type     |
| -------------  | -------------------------------------------  |-------------|
| brand_name     | Name of the phone brand                      |     chr     |
| model_name     | Name of the phone model                      |     chr     |
|    os          |    Operating system                          |     chr     |
| Popularity     | The popularity of the phone in range 1-1224  |     int     |
| Best Price     | Best price of the price-range in (UAH)       |     num     |
| Lowest Price   | Highest price of the price-range in (UAH)    |     num     |
| highest_price  | Lowest price of the price-range in (UAH)     |     num     |
| Sellers amount |   The amount sellers sold the phone          |     num     |
| Screen size    |  The size of phone's screen (inches).        |     num     |
| Memory size    |  The size of the phone's memory (GB)         |     num     |
| Battery size   |  The size of the phone's battery (mAh)       |     num     |
| Release date   |  The launch date of the product on the market|     chr     |


## Dataset cleaning

### Removing the index column
```{r}
df$X = NULL
```

### Factorizing the categorical variables

```{r}
df$brand_name <- factor(df$brand_name)
df$os <- factor(df$os)
```

### Managing missing variables
#### Battery_size and screen_size handling
We decided to drop the rows having missing variables for `battery_size` and `screen_size` since they were in low number.
```{r}
df <- df[- which(is.na(df$battery_size)),]
df <- df[- which(is.na(df$screen_size)),]
```

For the rows having missing values for `highest_price` and `lowest_price` we substitued them with a value that is resistant to outliers, namely the median.
#### Highest/Lowest_price handling
```{r}
tmp <- which(is.na(df$lowest_price))
df$lowest_price[tmp] <- median(df$lowest_price[-tmp])
```
```{r}
tmp <- which(is.na(df$highest_price))
df$highest_price[tmp] <- median(df$highest_price[-tmp])

```

#### Memory_size handling
We noticed that the only rows having `NA` in the `memory_size` feature were those with `os == ""`. Further investigation shows that this `blank` os refers to very simple phones that are usually given to elderly people. We decided to label them as `other`.
```{r}
# Replace empty strings in df$os with "other"
df$os <- sub("^$", "other", df$os)
```
```{r}
summary(df[df$os == "other", "memory_size" ])
```
```{r}
# Plot the standardized 'memory_size' column
memory_size_other <- df[df$os == "other", "memory_size"]

# Get the indices of the top 5 values
top5_indices <- order(memory_size_other, decreasing = TRUE)[1:5]

# Create a color vector
colors <- rep("black", length(memory_size_other))
colors[top5_indices] <- "red"

# Plot the 'memory_size' column for 'other' OS with the top 5 values colored red
plot(memory_size_other, main = "Memory Size for 'other' OS", ylab = "Memory Size", col = colors, type = "p")
```
Given the presence of such big outliers, we decided to substitute the NA with the median of the data having "other" as the `os` feature.
```{r}
tmp <- df[which(df$os == "other"),]$memory_size
df$memory_size[which(is.na(df$memory_size))] <- median(tmp[-which(is.na(tmp))])
```


### Converting prices

#### Aux function for € conversion
```{r}
# Function to convert currency A to Euros
convert_to_euros <- function(value, conversion_rate = 0.024) {
  # Return the value in Euros
  return(value * conversion_rate)
}
```
#### Conversion
```{r}
df$best_price <- convert_to_euros(df$best_price)
df$lowest_price <- convert_to_euros(df$lowest_price)
df$highest_price <- convert_to_euros(df$highest_price)
```

### Improving readibility of the os variable

```{r}
table(df$os)
```

We notice that, for the `os` feature, __2__ are  __EMUI__, __1__ is __KAIOS__, __3__ are __OxygenOS__ and __3__ are __WindowsPhone__. We decided to lable them either __Android__ or __other__ based on which class theyr features were closer to.


#### Labeling of the other OSs

```{r}
# Function to calculate the mean of all numeric columns for each value of df$os
calculate_means <- function(df) {
  # Split the data frame by the 'os' column
  split_df <- split(df, df$os)
  
  # Function to calculate the mean of each column
  mean_fun <- function(x) {
    if(is.numeric(x)) mean(x, na.rm = TRUE) else NULL
  }
  
  # Apply the function to each subset of the data frame
  mean_df <- lapply(split_df, function(x) sapply(x, mean_fun))
  
  # Remove NULL elements (non-numeric columns)
  mean_df <- lapply(mean_df, function(x) x[!sapply(x, is.null)])
  
  # Convert the list back to a data frame
  mean_df <- do.call(rbind, mean_df)
  
  return(mean_df)
}

calculate_means(df)
```

From these results we decided to relable them in this way:
 - __EMUI__ -> __Android__
 - __KAIOS__ -> __other__
 - __OxygenOS__ -> __Android__
 - __WindowsPhone__ -> __Android__

```{r}
# Recode the 'os' column
levels(df$os) <- c("other", "Android", "Android", "iOS", "other", "Android", "Android")
```

### Handling outlier

From our previous data exploration there was one phone having `os == "other"` with a relatively big value for its `memory_size` feature. We decided to lable it as __Android__.

```{r}
df[df$os == "other", ][which.max(df$memory_size[df$os == "other"]),]$os = "Android"
```

### Splitting the timestamp

To better explore the relationship between `best_price` and `release_date` we decided to split the latter variable in two features: `month` and `year`. For simpler reading we subtracted 2000 to the years.
```{r}
df$month <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][1]}))
df$year <- as.numeric(sapply(df$release_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][2]}))-2000
```

### Checking for duplicates

We first check for duplicates inside the data, not considering the features `popularity` and `model_name`
```{r}
idxs <- which(duplicated(df[,c(-2, -4)]))

df[idxs[0:10],]
```

We decide to aggregate the duplicates in one row, having its `popularity` subsituted by the mean of the duplicates.
```{r}
idxs <- which(duplicated(df[,c(-2, -4)]))

succ <- idxs[-length(idxs)] - idxs[-1] == -1
succ <- c(succ, FALSE)

i = 1
while (i <= length(idxs)){
  start = idxs[i]-1
  sum <- c(df$popularity[idxs[i]])
  while (succ[i] == TRUE){
    i = i + 1
    sum <- c(sum, df$popularity[idxs[i]])
  }
  df$popularity[start] <- mean(c(sum, df$popularity[start]))
  i = i + 1
}

df <- df[-idxs, ]
```

### Handling popularity

We decided to group phones by their `popularity` in a new variable called `popularity_level` defined as:
  - Phones below the first `popularity` quartile are labeled as __low__
  - Phones above the first `popularity` quartile and below the second are labeled as __low-medium__
  - Phones above the second `popularity` quartile and below the third are labeled as __medium-high__
  - Phones above the third `popularity` quartile are labeled as __high__

```{r}
df$popularity <- as.numeric(df$popularity)

tag <- quantile(df$popularity)

df$popularity_levels <- cut(df$popularity, breaks = tag,
labels=c("low", "medium-low", "medium-high", "high"), include.lowest=TRUE)
```

## Dataset exploration

### Correlation matrix
We begin by computing the correlation matrix between the quantitative variables

```{r}
# Select the numeric columns
df_numeric <- df[sapply(df, is.numeric)]

# Compute the correlation matrix
cor_matrix <- cor(df_numeric, use = "complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, lab = TRUE, colors=c("yellow", "white", "purple"))
```

From this matrix we can deduce:
  - The features `highest_price` and `lowest_price` are pretty much linearly correlated with `best_price`
  - `memory_size` also shows a strong positive correlation with `best_price` and a mild postive correlation with `screen_size`
  - `year`, `popularity`, `sellers_amount` and `screen_size` show a light positive correlation
  - `month` and `battery_size` show almost no linear correlation with `best_price`

### Barplots

For the quantitive variables we decided to plot the barplots. Notably we decided to color the bars based on their `os` feature, since we expect the phones with `os == "other"`
to have a different distribution than Android and iOs.

```{r}
cols <- c(5, 8:11)

for (i in cols) {
  layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
  par(mar=c(0,3.1,1.1,2.1))
  boxplot(df[, i], horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main=colnames(df)[i]) # xaxt
  par(mar=c(4,3.1,1.1,2.1))
  hist(df[, i],border = F, main = "", col="purple")
}
```

```{r}
  layout(mat=matrix(c(1,2),2,1,byrow = TRUE), height = c(1,8) )
  par(mar=c(0,3.1,1.1,2.1))
  boxplot(df[, "screen_size"], horizontal = TRUE,  xaxt = "n", col= "yellow", frame= F, main=colnames(df)["screen_size"]) # xaxt
  par(mar=c(4,3.1,1.1,2.1))
  hist(df[, "screen_size"],border = F, main = "", col="purple")
```

```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a stacked barplot of df$screen_size colored by df$os
ggplot(df, aes(x = factor(screen_size), fill = os)) +
  geom_bar(position = "stack") +
  theme_minimal() +
  labs(title = "Screen Size by OS", x = "Screen Size", y = "Count")
```

```{r}
# Create a histogram of df$screen_size colored by df$os
hist_screen_size <- ggplot(df, aes(x = screen_size, fill = os)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 12) +
  labs(title = "Histogram of Screen Size by OS", x = "Screen Size", y = "Count")

# Create a boxplot of df$screen_size colored by df$os
box_screen_size <- ggplot(df, aes(x = os, y = screen_size, fill = os)) +
  geom_boxplot() +
  labs(title = "Boxplot of Screen Size by OS", x = "OS", y = "Screen Size")

# Arrange the histogram and boxplot in a grid
grid.arrange(hist_screen_size, box_screen_size, ncol = 2)

# Create a histogram of df$memory_size colored by df$os
hist_screen_size <- ggplot(df, aes(x = memory_size, fill = os)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 10) +
  labs(title = "Histogram of Memory Size by OS", x = "Memory Size", y = "Count")

# Create a boxplot of df$memory_size colored by df$os
box_screen_size <- ggplot(df, aes(x = os, y = memory_size, fill = os)) +
  geom_boxplot() +
  labs(title = "Boxplot of Memory Size by OS", x = "OS", y = "Memory Size")

# Arrange the histogram and boxplot in a grid
grid.arrange(hist_screen_size, box_screen_size, ncol = 2)

# Create a histogram of df$battery_size colored by df$os
hist_screen_size <- ggplot(df, aes(x = battery_size, fill = os)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 12) +
  labs(title = "Histogram of Battery Size by OS", x = "Battery Size", y = "Count")

# Create a boxplot of df$battery_size colored by df$os
box_screen_size <- ggplot(df, aes(x = os, y = battery_size, fill = os)) +
  geom_boxplot() +
  labs(title = "Boxplot of Battery size by OS", x = "OS", y = "Battery Size")

# Arrange the histogram and boxplot in a grid
grid.arrange(hist_screen_size, box_screen_size, ncol = 2)

```